Equivalence Hypergraphs: E-Graphs for Monoidal Theories

### 2023-11-28 sections/combinatorial_semantics.tex
Refined parent and consistency relations WIP

    An e-hypergraph $\mathcal{G}$ over a monoidal signature $\Sigma$ is a tuple \[(V_{\mathcal{G}},E_{\mathcal{G}},s,t,l,<_c,\#)\] where $(V_{\mathcal{G}},E_{\mathcal{G}},s,t)$ are the same as in~\ref{def:hypergraph}.
    The labelling $l$ is modified to include an extra value $\bot \;$ $l :E_{\mathcal{G}} \to \Sigma + 1$. 
    % With an abuse of notation, we will confuse $\mathcal{G}$ with $V_{\mathcal{G}} \cup E_{\mathcal{G}}$, when treating it as a set.
    \question{Shall I use $\cup$ instead of $+$?}
    $<_c \subset (V_\mathcal{G} + E_\mathcal{G}) \times (V_\mathcal{G} + E_\mathcal{G})$ is a child relation imposing a strict partial order on $(V_\mathcal{G} + E_\mathcal{G})$.
    We require that for a given $x \in (V_\mathcal{G} + E_\mathcal{G})$ a set of all predecessors of $x\;$ $[x) = \{x' \in E_\mathcal{G} \;||\; x' <_c x\; \}$ is finite and consists of edges $e \in E_{\mathcal{G}}$ such that $l(e) = \bot$ only.
    Next, each $x$ has at most one immediate predecessor.
    \update{We will denote the fact that $e$ is an immediate predecessor of $v$ as $e <_c^{\mu} v$.}
    % , which essentially makes $<_{c}$ into a function that for any given $x$ returns its immediate predecessor.
    Such order was called a \textbf{well-founded arborescent} strict partial order in~\cite{orders_and_events}.
    Similarly, a set of all successors is $(x] = \{x' \in (V_\mathcal{G} + E_\mathcal{G}) \; ||\; x <_{c} x'\}$ where $x'$ can be a node or an edge. 
    % An edge $e \in E_{\mathcal{G}}$ is unlabelled, i.e. $l(e) = \bot$ iff $(e] \not = \varnothing$.
    Finally, we require that $<_c^{\mu}$ respects connectivity, i.e. if $v \in s(e)$ then $e' <_c^{\mu} e$ iff $e' <_c^{\mu} v$.

    $\langle (V_\mathcal{G} + E_\mathcal{G}), <_c, \# \rangle$ is an event-structure where $\# \subset (V_\mathcal{G} + E_\mathcal{G}) \times (V_\mathcal{G} + E_\mathcal{G})$ is an irreflexive symmetric relation called \textit{conflict} relation which satisfies the following
    \begin{enumerate}
        \label{def:e-hypergraph-req}
        \item 
        \[
    \forall x_1, x_2, x_3 \in (V_\mathcal{G} + E_\mathcal{G}) \text{, if } x_1 <_c x_2 \text{ and } x_1 \# x_3 \text{ then } x_2 \# x_3
    \] 
    if $x_1 <_c x_2$ we say that conflict $x_2 \# x_3$ is inherited from the conflict $x_1 \# x_3$. If the conflict is not inherited we say it is \textit{immediate} and denote it with $\#_{\mu}$
    \item $<_c$ and $\#$ are mutually exclusive
    \item $\#$ is defined \textit{only} for $x \in (V_\mathcal{G} + E_\mathcal{G})$ such that $[x) \not \eq \varnothing$
    \item $\#$ should respect connectivity
    \item \update{Complement of $\conflict_{\mu}$ should be transitive}
    \item for all distinct $x_1, x_2, x_3 \in (V_\mathcal{G} + E_\mathcal{G})$ $x_1 \#_{\mu} x_2$ and $x_2 \#_{\mu} x_3$ implies $x_1 \#_{\mu} x_3$
    \item for all $x, x' \in (V_\mathcal{G} + E_\mathcal{G})$ $x \#_{\mu} x'$ implies $[x) = [x')$
        \end{enumerate}

The last two requirements are known as \textit{confusion freeness}~\cite{ConfusionFreeEvents}.
\end{definition}

\Aleksei{Switch to the complent in the definition of the relation and call it a \textit{consistency} relation}

Intuitively, the child and conflict relation together make it possible to model e-classes in our e-hypergraphs: an unlabelled edge $e$ represents an e-class where the children of $e$ that are in immediate conflict correspond to nodes and edges in different e-class instances.
Condition (5) states that the following situation is not possible: $x_1 \# x_3$ and $x_1 \bar{\#} x_2$ and $x_2 \bar{\#} x_3$.
Transitivity of $\bar{\#}$ requires that in this case $x_1 \bar{\#} x_3$.

\begin{lemma}[Complement of $\#$ is equivalence relation]
\begin{proof}
As $\#$ is irreflexive then its complement $\bar{\#}$ is reflexive because $\forall x \; . \; (x,x) \not \in \#$. Because $\#$ is symmetric so is $\bar{\#}$, otherwise if $(x_1, x_2) \in \bar{\#}$ and $(x_2,x_1) \not \in \bar{\#}$ then $(x_2,x_1) \in {\#}$ which is symmetric and hence $(x_1,x_2)$ should be in ${\#}$.
Finally, the complement is transitive because of (5).

\end{proof}
\end{lemma}

@@

% We associate an e-class with a hierarchical edge, i.e. an edge with children. 
% The parent functions satisfy some conditions. First, an edge and any of its source and target vertices must have the same parent: $p_V (v) = p_E(e) = p_V (v')$ for all $v \in s(e)$ and $v' \in t(e)$, respectively.
% Second, the parent relation must be acyclic.
% More precisely, we assume for all $e \in E$ there is some $k \geq 1$ such that $(p_E,\bot)k(e) = \bot$ where $\bot$ is the element of $1$ and $p_{E,\bot} :E+1 \to E+1$ is the extension of $p_E$ adding $p_{E,\bot}(\bot)=\bot$.

@@

@@

        \item
    if $x_1 \#_{\mathcal{F}} x_2$ then $\phi(x_1) \#_{\mathcal{G}} \phi(x_2)$, where $x_1, x_2 \in \mathcal{F}$.
    Or, written functionally, when $x \in E_{\mathcal{F}}$
    \[
        \langle \phi_{V};i_{V_{\mathcal{G}}}, \phi_{E};i_{E_{\mathcal{G}}} \rangle^{*}(\conflict_{\mathcal{F}}(i_{E_{\mathcal{F}}}(x)))
        \subseteq
        \conflict_{\mathcal{G}}(\phi_{E};i_{E_{\mathcal{G}}}(x))
    \]
    , where $\phi_{V};i_{V_{\mathcal{G}}} : V_{\mathcal{F}} \to V_{\mathcal{G}} + E_{\mathcal{G}}$, and similarly for $\phi_{E};i_{E_\mathcal{G}}$ so that $\langle \phi_{V};i_{V_{\mathcal{G}}}, \phi_{E};i_{E_{\mathcal{G}}} \rangle : V_{\mathcal{F}} + E_{\mathcal{F}} \to  V_{\mathcal{G}} + E_{\mathcal{G}}$.
    And, when $x \in V_{\mathcal{F}}$
    \[
        \langle \phi_{V};i_{V_{\mathcal{G}}}, \phi_{E};i_{E_{\mathcal{G}}} \rangle^{*}(\conflict_{\mathcal{F}}(i_{V_{\mathcal{F}}}(x)))
        \subseteq
        \conflict_{\mathcal{G}}(\phi_{V};i_{V_{\mathcal{G}}}(x))
    \]
    Or, on per-component basis, if $i_{E_{\mathcal{F}}}(x_1) \in \conflict_{\mathcal{F}}(i_{E_{\mathcal{F}}}(x_2))$ then $i_{E_{\mathcal{G}}}(\phi_{E}(x_1)) \in \conflict_{\mathcal{G}}(i_{E_{\mathcal{G}}}(\phi_{E}(x_2)))$ and similarly for other types of $x_1,x_2$.
\end{enumerate}

@@

@@

    Then, the pushout of e-hypergraphs $X$ and $Y$ is computed in two steps.
    First, a coproduct of $X+Y$ is computed which is 
    \[
        X + Y = \{V_{X} + V_{Y}, E_{X} + E_{Y}, s_{X+Y}, t_{X+Y}, \#_{X+Y}, <_{c,X+Y} \}
    \]

@@

        \item $u$ has no pre-image in $Z$ and then 

@@

            <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}[u]) = [<_{c,X+Y}^{\mu}(i_{V_{X} + V_{Y}}(u))]

@@

@@

@@

        \item We have the only case of $u$ having no pre-image in $E_{Z}$ because its empty.

@@

                \item $[u) = \varnothing$ and for all $v_i \in s(u)$ and $v_j \in t(u)$ $[[v_i]) \not = \varnothing$ and $[[v_j]) \not = \varnothing$.

@@

                    \] where $v$ is either in $s([u])$ or $t([u])$.

@@

    \begin{remark}
        Essentially, what we build above is that $[e] = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]))$ when there exists $v' \in V_{X} + V_{Y}$ such that $[v'] = [v]$ and $e = <_{c,X+Y}^{\mu}(v')$.
    \end{remark}
    \update{If $f$ and $g$ are not mono, the definitions stay the same as $i_{2,E}$ and $i_{1,E}$ are mono.}
    Clearly, all the cases above are disjoint.

@@

    We also need to makes sure that $[]_{E} : E_{X} + E_{Y} \to \sfrac{E_{X} + E_{Y}}{\sim}$ and $[]_{V} : V_{X} + V_{Y} \to \sfrac{V_{X} + V_{Y}}{\sim}$ are homomorphisms.
    \begin{enumerate}
        \item $s_{X+Y};[]_{V}^{*}(e) = []_{E};s_{\sfrac{X+Y}{\sim}}(e)$
              This is the same as in $\catname{Hyp_{\Sigma}}$.
        \item $[<_{c,X+Y}^{\mu}(i_{E_{X} + E_{Y}}(e))]_{E} = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([e]_{E}))$ for $e$ such that $[e) \not = \varnothing$.
              Note that by requiring $[e) \not = \varnothing$ we use the first definition for edges.
              This is homomorphic by definition. 
        \item $[<_{c,X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v))]_{E} = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]_{V}))$.
              Let's make case analysis on $[v]_{V}$.
              \begin{itemize}
                \item $v$ has a pre-image $z$ in $V_{Z}$ such that $v = f_{V};i_{1,V}(z)$.
                    \begin{itemize}
                        \item If $[f_{V}(z)) = \varnothing$ and $[g_{V}(z)) \not = \varnothing$, then, because $i_{1,V}$ is a coproduct injection, it must be the case that $[i_{1,V}(f_{V}(z))) = \varnothing$ and we do not require any equalities for such $v$.
                        \item If $[f_{V}(z)) \not = \varnothing$ and $[g_{V}(z)) = \varnothing$, then,
                              by definition, we have $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{V_{X} + V_{Y}}[v]) = [f_{V};i_{V_{X}};<_{c,X}^{\mu};i_{1,E}(z)] = [i_{1,E}(<_{c,X}^{\mu}(f_{V};i_{V_{X}}(z)))] = [i_{1,E}(e_x)]$.
                              Because $i_{1,E}$ is a homomorphism, we also have $[<_{c,X+Y}^{\mu}(f_{V};i_{1,V};i_{V_{X} + V_{Y}}(z))] = [i_{1,E}(<_{c,X}^{\mu}(f_{V};i_{V_{X}}(z)))] = [i_{1,E}(e_x)]$.
                    \end{itemize}
                \item The case when $v = g_{V};i_{2,V}(z)$ is symmetric.
                \item If $v$ has no pre-image in $V_{Z}$ then homomorphism-ness of $[]_{V}$ is similar to the case for $[]_{E}$ above.
              \end{itemize}
    \end{enumerate}

    Then $<_{c,\sfrac{X+Y}{\sim}} : V_{\sfrac{X+Y}{\sim}} + E_{\sfrac{X+Y}{\sim}} \to (E_{\sfrac{X+Y}{\sim}})^{*}$ is defined as a transitive closure of $<_{c,\sfrac{X+Y}{\sim}}^{\mu}$.
    \question{I probably should've added injections for $<$ as well as it is defined on a coproduct (like I do below for $\conflict$) }

    \update{Check that $<_{c,\sfrac{X+Y}{\sim}}^{\mu}$ is irreflexive and assymetric and that it respects connectivity.}

>>>>

    An e-hypergraph $\mathcal{G}$ over a monoidal signature $\Sigma$ is a tuple \[(V_{\mathcal{G}},E_{\mathcal{G}},s,t,l,<_c,\#)\] where $(V_{\mathcal{G}},E_{\mathcal{G}},s,t)$ are the same as in~\ref{def:hypergraph}.
    The labelling $l$ is modified to include an extra value $\bot \;$ $l :E_{\mathcal{G}} \to \Sigma + 1$. 
    % With an abuse of notation, we will confuse $\mathcal{G}$ with $V_{\mathcal{G}} \cup E_{\mathcal{G}}$, when treating it as a set.
    \question{Shall I use $\cup$ instead of $+$?}
    $<_c \subset (V_\mathcal{G} + E_\mathcal{G}) \times (V_\mathcal{G} + E_\mathcal{G})$ is a child relation imposing a strict partial order on $(V_\mathcal{G} + E_\mathcal{G})$.
    We require that for a given $x \in (V_\mathcal{G} + E_\mathcal{G})$ a set of all predecessors of $x\;$ $[x) = \{x' \in E_\mathcal{G} \;||\; x' <_c x\; \}$ is finite and consists of edges $e \in E_{\mathcal{G}}$ such that $l(e) = \bot$ only.
    Next, each $x$ has at most one immediate predecessor.
    \update{We will denote the fact that $e$ is an immediate predecessor of $v$ as $e <_c^{\mu} v$.}
    Similarly, a set of all successors is $(x] = \{x' \in (V_\mathcal{G} + E_\mathcal{G}) \; ||\; x <_{c} x'\}$ where $x'$ can be a node or an edge. 
    % An edge $e \in E_{\mathcal{G}}$ is unlabelled, i.e. $l(e) = \bot$ iff $(e] \not = \varnothing$.
    Finally, we require that $<_c^{\mu}$ respects connectivity, i.e. if $v \in s(e)$ then $e' <_c^{\mu} e$ iff $e' <_c^{\mu} v$.

    Then, consider a relation $\consistency^{\mu} \subseteq (V_{\mathcal{G}} + E_{\mathcal{G}}) \times (V_{\mathcal{G}} + E_{\mathcal{G}})$ which we will call a \textit{consistency} relation and which has the following properties.
    \begin{enumerate}
        \label{def:consistency_properties}
        \item For each $v \in V_{\mathcal{G}}$ $v \consistency^{\mu} v$
        \item For each $e \in E_{\mathcal{G}}$ $e \consistency^{\mu} e$
        \item If $v \in s(e)$ then $v \consistency^{\mu} e$
        \item If $v \in t(e)$ then $v \consistency^{\mu} e$
        \item The relation is symmetric and transitive
        \item The relation is defined only for edges $e$ and nodes $v$ such that $[v) = [e) \not = \varnothing$
    \end{enumerate}

%     $\langle (V_\mathcal{G} + E_\mathcal{G}), <_c, \# \rangle$ is an event-structure where $\# \subset (V_\mathcal{G} + E_\mathcal{G}) \times (V_\mathcal{G} + E_\mathcal{G})$ is an irreflexive symmetric relation called \textit{conflict} relation which satisfies the following
%     \begin{enumerate}
%         \label{def:e-hypergraph-req}
%         \item 
%         \[
%     \forall x_1, x_2, x_3 \in (V_\mathcal{G} + E_\mathcal{G}) \text{, if } x_1 <_c x_2 \text{ and } x_1 \# x_3 \text{ then } x_2 \# x_3
%     \] 
%     if $x_1 <_c x_2$ we say that conflict $x_2 \# x_3$ is inherited from the conflict $x_1 \# x_3$. If the conflict is not inherited we say it is \textit{immediate} and denote it with $\#_{\mu}$
%     \item $<_c$ and $\#$ are mutually exclusive
%     \item $\#$ is defined \textit{only} for $x \in (V_\mathcal{G} + E_\mathcal{G})$ such that $[x) \not \eq \varnothing$
%     \item $\#$ should respect connectivity
%     \item \update{Complement of $\conflict_{\mu}$ should be transitive}
%     \item for all distinct $x_1, x_2, x_3 \in (V_\mathcal{G} + E_\mathcal{G})$ $x_1 \#_{\mu} x_2$ and $x_2 \#_{\mu} x_3$ implies $x_1 \#_{\mu} x_3$
%     \item for all $x, x' \in (V_\mathcal{G} + E_\mathcal{G})$ $x \#_{\mu} x'$ implies $[x) = [x')$
%         \end{enumerate}

% The last two requirements are known as \textit{confusion freeness}~\cite{ConfusionFreeEvents}.
\end{definition}

% Intuitively, the child and conflict relation together make it possible to model e-classes in our e-hypergraphs: an unlabelled edge $e$ represents an e-class where the children of $e$ that are in immediate conflict correspond to nodes and edges in different e-class instances.
% Condition (5) states that the following situation is not possible: $x_1 \# x_3$ and $x_1 \bar{\#} x_2$ and $x_2 \bar{\#} x_3$.
% Transitivity of $\bar{\#}$ requires that in this case $x_1 \bar{\#} x_3$.

% \begin{lemma}[Complement of $\#$ is equivalence relation]
% \begin{proof}
% As $\#$ is irreflexive then its complement $\bar{\#}$ is reflexive because $\forall x \; . \; (x,x) \not \in \#$. Because $\#$ is symmetric so is $\bar{\#}$, otherwise if $(x_1, x_2) \in \bar{\#}$ and $(x_2,x_1) \not \in \bar{\#}$ then $(x_2,x_1) \in {\#}$ which is symmetric and hence $(x_1,x_2)$ should be in ${\#}$.
% Finally, the complement is transitive because of (5).

% \end{proof}
% \end{lemma}

@@

\begin{remark}
    Because $\consistency^{\mu}$ respects connectivity and is reflexive, symmetric and transitive, it is essentially a relation between connected components that have the same predecessor.
    This notion will be helpful in the following example.
\end{remark}
% We associate an e-class with a hierarchical edge, i.e. an edge with children. 
% The parent functions satisfy some conditions. First, an edge and any of its source and target vertices must have the same parent: $p_V (v) = p_E(e) = p_V (v')$ for all $v \in s(e)$ and $v' \in t(e)$, respectively.
% Second, the parent relation must be acyclic.
% More precisely, we assume for all $e \in E$ there is some $k \geq 1$ such that $(p_E,\bot)k(e) = \bot$ where $\bot$ is the element of $1$ and $p_{E,\bot} :E+1 \to E+1$ is the extension of $p_E$ adding $p_{E,\bot}(\bot)=\bot$.

@@

A few examples of e-hypergraphs (and their corresponding graphical interpretation) can be seen in Figure~\ref{fig:e-hypergraph-example}.
In example~\ref{fig:example_a} the relations $<_{c}$ and $\consistency^{\mu}$ are the following
\[
    <_{c} = \{(v_3,e_1), (e_2, e_1), (v_4, e_1) \}
\] and 
\begin{align*}
    \consistency^{\mu} = \{ (v_1,e_1),
                            (v_2, e_1), 
                            (v_3,e_2), (v_4,e_2)\}^{c}
\end{align*}
where $c$ denotes the reflexive, syymetric and transitive closure.
In other words, $e_1$ is the predecessor of the connected component that contains edge $e_2$ and this connected component is consistent with itself only.
The connected component that contains $e_1$ is also consistent with itself only.

In the next example~\ref{fig:example_b} $e_1$ is the predecessor of two connected components but these components are consistent only with themselves.
We denote this fact by using a vertical line. 
Practically, hierarchical edges such as $e_1$ represent e-classes and connected components which are pairwise not consistent belong to different instances of the same e-class.

In the last example~\ref{fig:example_c} two connected components that contain $e_2$ and $e_3$ respectively are consistent with themselves and with each other.
\end{example}

\begin{figure}[t]
    \begin{subfigure}{0.3\linewidth}
        \[
            \scalebox{0.6}{
                \tikzfig{combinatorial_semantics/e-hypergraph-example-1}
            }
            \]
        \caption{}
        \label{fig:example_a}
    \end{subfigure}
    \hfill
    \begin{subfigure}{0.3\linewidth}
        \[
          \scalebox{0.6}{
            \tikzfig{combinatorial_semantics/e-hypergraph-example-2}
          }  
        \]
        \caption{}
        \label{fig:example_b}
    \end{subfigure}
    \hfill
    \begin{subfigure}{0.3\linewidth}
        \[
          \scalebox{0.6}{
            \tikzfig{combinatorial_semantics/e-hypergraph-example-3}
          }  
        \]
        \caption{}
        \label{fig:example_c}
    \end{subfigure}
    \caption{E-hypergraph examples}
    \label{fig:e-hypergraph-example}
\end{figure}

@@

        \item
    if $x_1 \consistency_{\mathcal{F}}^{\mu} x_2$ then $\phi(x_1) \consistency_{\mathcal{G}}^{\mu} \phi(x_2)$, where $x_1, x_2 \in \mathcal{F}$.
    Or, written functionally, when $x \in E_{\mathcal{F}}$
    \[
        \langle \phi_{V};i_{V_{\mathcal{G}}}, \phi_{E};i_{E_{\mathcal{G}}} \rangle^{*}(\consistency_{\mathcal{F}}^{\mu}(i_{E_{\mathcal{F}}}(x)))
        \subseteq
        \consistency_{\mathcal{G}}^{\mu}(\phi_{E};i_{E_{\mathcal{G}}}(x))
    \]
    , where $\phi_{V};i_{V_{\mathcal{G}}} : V_{\mathcal{F}} \to V_{\mathcal{G}} + E_{\mathcal{G}}$, and similarly for $\phi_{E};i_{E_\mathcal{G}}$ so that $\langle \phi_{V};i_{V_{\mathcal{G}}}, \phi_{E};i_{E_{\mathcal{G}}} \rangle : V_{\mathcal{F}} + E_{\mathcal{F}} \to  V_{\mathcal{G}} + E_{\mathcal{G}}$.
    And, when $x \in V_{\mathcal{F}}$
    \[
        \langle \phi_{V};i_{V_{\mathcal{G}}}, \phi_{E};i_{E_{\mathcal{G}}} \rangle^{*}(\consistency_{\mathcal{F}}^{\mu}(i_{V_{\mathcal{F}}}(x)))
        \subseteq
        \consistency_{\mathcal{G}}^{\mu}(\phi_{V};i_{V_{\mathcal{G}}}(x))
    \]
    % Or, on per-component basis, if $i_{E_{\mathcal{F}}}(x_1) \in \conflict_{\mathcal{F}}(i_{E_{\mathcal{F}}}(x_2))$ then $i_{E_{\mathcal{G}}}(\phi_{E}(x_1)) \in \conflict_{\mathcal{G}}(i_{E_{\mathcal{G}}}(\phi_{E}(x_2)))$ and similarly for other types of $x_1,x_2$.
\end{enumerate}

@@

\begin{definition}[Pushout in $\catname{EHyp_{\Sigma}}$]

@@

    Then, the pushout of e-hypergraphs $X$ and $Y$ is computed in two steps.
    First, a coproduct of $X+Y$ is computed which is 
    \[
        X + Y = \{V_{X} + V_{Y}, E_{X} + E_{Y}, s_{X+Y}, t_{X+Y}, \consistency_{X+Y}, <_{c,X+Y} \}
    \]

@@

        \item $u$ has no pre-image in $Z$ and $<_{c,X+Y}^{\mu}(i_{V_{X} + V_{Y}}(u)) = e'$

@@

            <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}[u]) = [<_{c,X+Y}^{\mu}(i_{V_{X} + V_{Y}}(u))] = [e']

@@

        \item $u$ has no pre-image in $Z$ and $<_{c,X+Y}^{\mu}(i_{V_{X} + V_{Y}}(u))$ is undefined (i.e. $[u) = \varnothing$), but there exists $[e]$ such that $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])) = [e']$ and $[u] \in t([e])$ or $[u] \in s([e])$, then
                <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}[u]) = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}[e])
              \update{Assumptions above guarantee that the predecessor of such node does not depend on the choice of $e$ if there are several of them.}

@@

    We have the only case of $u$ having no pre-image in $E_{Z}$ because it is empty.

@@

@@

                \item $[u) = \varnothing$ and for some $[v_i] \in s([u])$ and $[v_j] \in t([u])$ $[[v_i]) \not = \varnothing$ and $[[v_j]) \not = \varnothing$.

@@

                    \] where $[v]$ is either in $s([u])$ or $t([u])$.

@@

    \update{We will also build this relation in steps, first we do steps for nodes (1) - (3), then steps (1) - (2) for edges and finally, step (4) for nodes.}

    \begin{remark}
        Essentially, what we build above is that if the relation is undefined, it is inherited from either incident nodes or edges.
    \end{remark}
    \update{If $f$ and $g$ are not mono, the definitions stay the same as $i_{2,E}$ and $i_{1,E}$ are mono.}
    Clearly, all the cases above are disjoint.

@@

    We also need to makes sure that $[]_{E} : E_{X} + E_{Y} \to \sfrac{E_{X} + E_{Y}}{\sim}$ and $[]_{V} : V_{X} + V_{Y} \to \sfrac{V_{X} + V_{Y}}{\sim}$ are homomorphisms.
    \begin{enumerate}
        \item $s_{X+Y};[]_{V}^{*}(e) = []_{E};s_{\sfrac{X+Y}{\sim}}(e)$
              This is the same as in $\catname{Hyp_{\Sigma}}$.
        \item $[<_{c,X+Y}^{\mu}(i_{E_{X} + E_{Y}}(e))]_{E} = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([e]_{E}))$ for $e$ such that $[e) \not = \varnothing$.
              Note that by requiring $[e) \not = \varnothing$ we use the first definition for edges.
              This is homomorphic by definition. 
        \item $[<_{c,X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v))]_{E} = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]_{V}))$.
              Let's make case analysis on $[v]_{V}$.
              \begin{itemize}
                \item $v$ has a pre-image $z$ in $V_{Z}$ such that $v = f_{V};i_{1,V}(z)$.
                    \begin{itemize}
                        \item If $[f_{V}(z)) = \varnothing$ and $[g_{V}(z)) \not = \varnothing$, then, because $i_{1,V}$ is a coproduct injection, it must be the case that $[i_{1,V}(f_{V}(z))) = \varnothing$ and we do not require any equalities for such $v$.
                        \item If $[f_{V}(z)) \not = \varnothing$ and $[g_{V}(z)) = \varnothing$, then,
                              by definition, we have $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{V_{X} + V_{Y}}[v]) = [f_{V};i_{V_{X}};<_{c,X}^{\mu};i_{1,E}(z)] = [i_{1,E}(<_{c,X}^{\mu}(f_{V};i_{V_{X}}(z)))] = [i_{1,E}(e_x)]$.
                              Because $i_{1,E}$ is a homomorphism, we also have $[<_{c,X+Y}^{\mu}(f_{V};i_{1,V};i_{V_{X} + V_{Y}}(z))] = [i_{1,E}(<_{c,X}^{\mu}(f_{V};i_{V_{X}}(z)))] = [i_{1,E}(e_x)]$.
                    \end{itemize}
                \item The case when $v = g_{V};i_{2,V}(z)$ is symmetric.
                \item If $v$ has no pre-image in $V_{Z}$ then homomorphism-ness of $[]_{V}$ is similar to the case for $[]_{E}$ above.
              \end{itemize}
    \end{enumerate}
    Then $<_{c,\sfrac{X+Y}{\sim}} : V_{\sfrac{X+Y}{\sim}} + E_{\sfrac{X+Y}{\sim}} \to (E_{\sfrac{X+Y}{\sim}})^{*}$ is defined as a transitive closure of $<_{c,\sfrac{X+Y}{\sim}}^{\mu}$.

    Let's now define $\consistency_{\sfrac{X+Y}{\sim}}^{\mu}$.
    We can consider the original consistency relation as a function $\consistency_{X+Y}^{\mu} : (V_{X} + V_{Y}) + (E_{X} + E_{Y}) \to 2^{(V_{X} + V_{Y}) + (E_{X} + E_{Y})}$.
    Quotienting the range of the function gives us $\consistency_{X+Y}'^{\mu} : V_{X} + V_{Y} + E_{X} + E_{Y} \to 2^{(\sfrac{V_{X} + V_{Y}}{\sim}) + \sfrac{(E_{X} + E_{Y})}{\sim}}$, which is essentially $\langle []_{V};i_{\sfrac{V_{X} + V_{Y}}{\sim}}, []_{E};i_{\sfrac{E_{X} + E_{Y}}{\sim}} \rangle^{*}$ (a copairing extended to sequences that we will further denote as $\langle []_{V}^{\consistency^{\mu}} []_{E}^{\consistency^{\mu}} \rangle$) applied to the return value of $\consistency_{X+Y}^{\mu}$.
    We then define the new relation for edges as
    \begin{enumerate}
        \item If $\consistency_{X+Y}^{\mu}(i_{E_{X} + E_{Y}}(e)) \not = \varnothing$, then
                \begin{align*}
                    \consistency_{\sfrac{X+Y}{\sim}}^{\mu}(i_{E_{X} + E_{Y}}([e]_{E})) =
                    (\langle []_{V}^{\consistency}, []_{E}^{\consistency}\rangle^{*}(\consistency_{X+Y}^{\mu}(i_{E_{X} + E_{Y}}(e))))^{c}
                \end{align*}
        \item $\consistency_{X+Y}^{\mu}(i_{E_{X} + E_{Y}}(e)) = \varnothing$ and there exists $[v]$ in $s([e])$ or $t([e])$ such that $\consistency_{\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]) \not = \varnothing$, then
        \[
            \consistency_{\sfrac{X+Y}{\sim}}^{\mu}(i_{E_{X} + E_{Y}}([e]_{E})) = \consistency_{\sfrac{X+Y}{\sim}}^{\mu}(i_{V_{X} + V_{Y}}([v]_{V}))
        \]
    \end{enumerate}

    and for nodes
    \begin{enumerate}

    \item If there exists $v' \in V_{X+Y}$ such that $\consistency_{X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v')) \not = \varnothing$ and $v' \sim v$
    \[
        \consistency_{\sfrac{X+Y}{\sim}}^{\mu}(i_{V_{X} + V_{Y}}([v]_{V})) = (\langle []_{V}^{\consistency}, []_{E}^{\consistency} \rangle^{*}(\bigcup_{v' \in V_{X} + V_{Y} | [v']_{V} = [v]_{V}} \consistency_{X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v'))))^{c}
    \]
    \item If for all $v' \in V_{X+Y}$ such that $v' \sim v$, $\consistency_{X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v')) = \varnothing$ and there exists $e \in E_{X+Y}$ such that either $[v] \in s([e])$ or $[v] \in t([e])$ and $\consistency_{\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])) \not = \varnothing$, then
    \[
        \consistency_{\sfrac{X+Y}{\sim}}^{\mu}(i_{V_{X} + V_{Y}}([v]_{V})) = \consistency_{\sfrac{X+Y}{\sim}}^{\mu}(i_{E_{X} + E_{Y}}([e]_{E}))
    \]
    \end{enumerate}
        where $i_*$ are injections into $\sfrac{V_{X} + V_{Y}}{\sim} + \sfrac{E_{X} + E_{Y}}{\sim}$ and $c$ is a reflexive, symmetric and transitive closure.
        % That is, the relation above is the smallest relation that contains $\langle []_{V}^{\conflict}, []_{E}^{\conflict} \rangle^{*}(\bigcup_{x' \in E_{X} + E_{Y} | [x']_{E} = [x]_{E}} i_{E_{X} + E_{Y}}(x'))$ and such that if $[x_1] <_{c,\sfrac{X+Y}{\sim}}([x_2])$ and $i_{\sfrac{E_{X} + E_{Y}}{\sim}}[x_1] \in \conflict_{\sfrac{X+Y}{\sim}}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}[x_3])$ then $i_{\sfrac{E_{X} + E_{Y}}{\sim}}[x_2] \in \conflict_{\sfrac{X+Y}{\sim}}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}[x_3])$.
        % And similarly for nodes.

    Again, we need to check that this is a well-defined function and that our $([]_{V},[]_{E})$ is indeed a homomorphism with respect to $\consistency^{\mu}$.

    \begin{itemize}
        \item 
        \begin{itemize}
